Procrastination and Dementia

### 2025-04-29 manuscript/manuscript.qmd
Re rendering site :fire:

In a stationary Markov model, transition probabilities $p_{ij}$ are assumed constant over time. However, cognitive transitions may vary with age, time since diagnosis, or other covariates. To account for this, we generalize to a non-stationary Markov model, where $p_{ij}^{(t)}$ depends on covariates $\boldsymbol{x}^{(t)}$

$$
p_{ij}^{(t)} = P(X_{t+1} = j \vert X_t = i, \boldsymbol{x}^{(t)})
$$

@@

We model these time-varying transition probabilities using multinomial logistic regression. For a system with $K = 3$ states (using state $K$ as reference), the log-odds of transitioning to state $j$ relative to state $K$ are linear functions of covariates $\boldsymbol{x}(t)$:

$$
log \left( \frac{p_{ij}^{(t)}}{p_{ik}^{(t)}} \right) = \alpha_j^{(t)} + \beta_j^{T \;(t)} \boldsymbol{x}^{(t)} \qquad \text{for } j = 1, \dots K-1
$$
where, $\alpha_j^{(t)}$ is the intercept for state $j$ at time $t$, $\beta_j^{(t)}$ is the vector of regression coefficients for state $j$ at time $t$, and $\boldsymbol{x}^{(t)}$ is the vector of covariates at time $t$. The model estimates the transition probabilities as a function of covariates, allowing us to assess how procrastination and other factors influence cognitive transitions.

@@

In our sample of $903$ older American adults aged 50 years or older, $63.01\%$ were female $(n = 569)$, with a mean sample age of $67.73$ years $(\pm 9.84)$. At baseline, $84.81\%$ of the sample was classified as having normal cognition, $14.96\%$ as having MCI, and $1.35\%$ as having dementia.

The mean procrastination score was $36.73$ $(\pm 8.23)$, with a range of $12 - 60$. The sample characteristics are presented in Table \ref{tab:sample-characteristics}.

>>>>

In a stationary Markov model, transition probabilities $p_{ij}$ are assumed constant over time. However, cognitive transitions may vary with age, time since diagnosis, or other covariates. To account for this, we generalize to a non-stationary Markov model, where $p_{ij}^{(t)}$ depends on covariates $\boldsymbol{X}^{(t)}$

$$
p_{ij}^{(t)} = P(X_{t+1} = j \vert X_t = i, \boldsymbol{X}^{(t)})
$$

@@

We model these time-varying transition probabilities using multinomial logistic regression. For a system with $K = 3$ states (using state $K$ as reference), the log-odds of transitioning to state $j$ relative to state $K$ are linear functions of covariates $\boldsymbol{X}(t)$:

$$
log \left( \frac{p_{ij}^{(t)}}{p_{ik}^{(t)}} \right) = \alpha_j^{(t)} + \beta_j^{T \;(t)} \boldsymbol{X}^{(t)} \qquad \text{for } j = 1, \dots K-1
$$
where, $\alpha_j^{(t)}$ is the intercept for state $j$ at time $t$, $\beta_j^{(t)}$ is the vector of regression coefficients for state $j$ at time $t$, and $\boldsymbol{X}^{(t)}$ is the vector of covariates at time $t$. The model estimates the transition probabilities as a function of covariates, allowing us to assess how procrastination and other factors influence cognitive transitions.

@@

In our sample of $903$ older American adults aged 50 years or older, $63.01\%$ were female $(n = 569)$, with a mean sample age of $67.73$ years $(\pm 9.84)$. At baseline, $84.81\%$ of the sample was classified as having normal cognition, $14.96\%$ as having MCI, and $1.35\%$ as having dementia. A
